\documentstyle{article}

\input{header.latex}

\title{Vehicle Design in \xtank}
\author{Terry Donahue \and Dan Schmidt}
\date{\today}

\begin{document}

\maketitle

\section{Overview}

Designing a vehicle requires some thought.  You will need to make
several choices about the type of body, engine, armor, weapons, and
special abilities that your vehicle will have.  Each choice that you
make will affect various attributes of the vehicle.  These attributes,
such as cost, weight, engine power, and drag value, will in
turn determine how your vehicle performs in combat.

\subsection*{Selected Characteristics}

\begin{description}
  \item [Body:]        The external shape of your vehicle
  \item [Engine:]      The driving force inside the vehicle
  \item [Weapons:]     Unsubtle ways to defeat your opponents
  \item [Armor:]       Protection from enemy attacks
  \item [Specials:]    Subtle ways to defeat your opponents
  \item [Heat Sinks:]  Remove heat created by firing weapons
  \item [Suspension:]  Keeps the vehicle stable during turns
  \item [Treads:]      Increases your traction
  \item [Bumpers:]     Reduces damage from collisions
\end{description}

What you choose for these characteristics will in turn determine your
vehicle's calculated characteristics, such as maximum speed and turning
rate.  Formulas for these characteristics can be found in {\bf Formulas}.

\subsection*{The Interface}

The interface to \xtank's vehicle design module is very intuitive.
You simply click on the menu entry corresponding to the part you would
like to change.  When you are asked to type in a response, type it and
then press Return.

\newpage

\section{Bodies}

Choosing a body is the most important decision to make when designing
a vehicle.  It determines the basic attributes of the vehicle.

\begin{description}
  \item [Size:]  The general size of your vehicle.  Generally, the bigger
your vehicle, the more costly and heavy its accessories will be.
  \item [Weight:]  How much the body weighs.
  \item [Weight Limit:]  How much weight the body can carry.
  \item [Space Limit:]  How much space is available inside the body for
things like weapons and engines.
  \item [Drag Value:]  How aerodynamic the body is, used to determine maximum
speed.  The higher the value, the better.
  \item [Handling:]  How maneuverable the body is, used to determine
things like your maximum turning rate.  See {\bf Suspension}.
  \item [Turrets:]  The number of turrets the body has to fire weapons from.
  \item [Cost:]  The basic cost of this vehicle.
\end{description}

\begin{center}
{\small
\begin{tabular}{|l|r|r|r|r|r|r|r|r|}                                    \hline
\multicolumn{9}{|c|}{\huge Bodies}                                  \\ \hline
     &      &        & Weight & Space & Drag  &          &         &    \\
Body & Size & Weight & Limit  & Limit & Value & Handling & Turrets & Cost
\\ \hline \hline
Lightcycle & 2 & 200 Kg & 800 Kg & 600 L & 0.10 & 8 & 0 & 3000 
\\ \hline
Hexo & 3 & 1500 Kg & 5000 Kg & 4000 L & 0.25 & 6 & 1 & 4000            
\\ \hline
Spider & 3 & 2500 Kg & 8000 Kg & 3000 L & 0.40 & 7 & 1 & 5000          
\\ \hline
Psycho & 4 & 5000 Kg & 18000 Kg & 8000 L & 0.60 & 4 & 1 & 5000         
\\ \hline
Tornado & 4 & 7000 Kg & 22000 Kg & 12000 L & 0.80 & 4 & 1 & 7000       
\\ \hline
Marauder & 5 & 9000 Kg & 28000 Kg & 18000 L & 1.00 & 3 & 2 & 10000     
\\ \hline
Tiger & 6 & 11000 Kg & 35000 Kg & 22000 L & 1.50 & 5 & 1 & 12000       
\\ \hline
Rhino & 7 & 12000 Kg & 40000 Kg & 30000 L & 2.00 & 3 & 2 & 10000       
\\ \hline
Medusa & 7 & 14000 Kg & 40000 Kg & 25000 L & 1.20 & 4 & 3 & 15000      
\\ \hline
Malice & 5 & 4000 Kg & 20000 Kg & 15000 L & .40 & 7 & 1 & 17000        
\\ \hline
\end{tabular}
}
\end{center}

\newpage

\section{Engines}

\subsection*{Normal Engines}
\begin{description}
  \item [Electric:]  Very small, super light, very low power. 
                     Good for motorcycles and small tanks.
  \item [Internal Combustion:]  Small, light, low power.      
                                Good for small tanks.
  \item [Turbine:]  Average size, weight and power.           
                    Good for all tanks.
\end{description}

\subsection*{Esoteric Engines}

\begin{description}
  \item [Fuel Cell:]  Very small, high power.
                      Good for small but fast tanks.
  \item [Fission:]  Huge, high power and long time between refills.
                    Good for big tanks.
  \item [Breeder Fission:]  huger, stronger, version of fission engine.
                            Good for big fast tanks.
  \item [Fusion:]  Ultimate powerhouse, very heavy, stays on the 
                   road for a while.  Good for behemoth tanks.
\end{description}

\subsection*{What the Numbers Mean}
\begin{description}
  \item [Power:]  Affects maximum speed and acceleration.
  \item [Weight:] How much the engine weighs.
  \item [Space:]  How much space the engine takes up.
  \item [Fuel Type:] Means nothing in the game; included for the curious.
  \item [Fuel Cost:] How much each liter of fuel will cost you when you
                     refuel during a battle.
  \item [Capacity:] How much fuel the engine can hold.
  \item [Cost:]  How much the engine costs.  Comes with a free full tank.
\end{description}
	
\begin{center}
{\small
\begin{tabular}{|l|r|r|r|l|r|r|r|}                                   \hline
\multicolumn{8}{|c|}{\huge Engines}                            \\ \hline
 & & & & & Fuel & &                                 \\
Engine & Power & Weight & Space & 
   Fuel Type & Cost & Capacity & Cost               \\ \hline \hline
Small Electric & 50 hp & 100 Kg & 20 L & Electricity & 5 & 200 & 1500
\\ \hline
Medium Electric & 100 hp & 150 Kg & 30 L & Electricity & 5 & 300 & 2200
\\ \hline
Large Electric & 200 hp & 200 Kg & 40 L & Electricity & 5 & 400 & 3000
\\ \hline
Super Electric & 300 hp & 250 Kg & 50 L & Electricity & 5 & 500 & 6000
\\ \hline
Small Combustion & 300 hp & 400 Kg & 200 L & Gasoline & 8 & 200 & 2000
\\ \hline
Medium Combustion & 400 hp & 500 Kg & 300 L & Gasoline & 8 & 300 & 2500
\\ \hline
Large Combustion & 500 hp & 600 Kg & 400 L & Gasoline & 8 & 400 & 3000
\\ \hline
Super Combustion & 600 hp & 1000 Kg & 600 L & Gasoline & 8 & 500 & 4000
\\ \hline
Small Turbine & 600 hp & 1000 Kg & 800 L & Kerosene & 10 & 350 & 4000   
\\ \hline
Medium Turbine & 700 hp & 1200 Kg & 1000 L & Kerosene & 10 & 450 & 5000 
\\ \hline
Large Turbine & 800 hp & 1500 Kg & 1500 L & Kerosene & 10 & 550 & 7000  
\\ \hline
Turbojet Turbine & 1000 hp & 2000 Kg & 2000 L & Kerosene & 10 & 750 & 10000 
\\ \hline
Fuel Cell & 1200 hp & 1000 Kg & 400 L & Hydrogen & 20 & 600 & 15000         
\\ \hline
Fission & 1500 hp & 3000 Kg & 3500 L & Uranium & 15 & 1000 & 20000         
\\ \hline
Breeder Fission & 1800 hp & 4000 Kg & 4000 L & Plutonium & 15 & 1250 & 25000
\\ \hline
Fusion & 2250 hp & 4000 Kg & 2500 L & Deuterium & 5 & 1500 & 40000         
\\ \hline
\end{tabular}
}
\end{center}

\newpage

\section{Weapons}

\subsection*{Types of Weapons}

\begin{description}
  \item [Machine Guns:]  Short range, high firing rate.  Low cost way to
do damage.
  \item [Pulse Rifles:]  Medium-long range, slower firing rate.  More expensive,
but wins out in long range battles.
  \item [Autocannons:]  High damage, medium range, high firing rate.  Best
qualities of first two; priced to match.
  \item [Rocket Launchers:]  Super damage, long range, slow firing rate.
Nastiest weapon, but not much ammo.
  \item [Acid Sprayer:]  High damage, very short range, high firing rate.
Great to use after running into an enemy.
  \item [Flame Thrower:] Low damage, super firing rate, short range.
Good against lightly armored enemies.
  \item [Mine Layer:] High damage, the only way to damage a vehicle's bottom
armor.
  \item [Heat Seeker:] High damage, turn in midair to track down
enemies.  The only way to damage a vehicle's top armor.  Very
expensive, very large.  
  \item [Oil Slick:] No damage, but vehicles driving through it lose control.
\end{description}

\subsection*{What the Numbers Mean}

\begin {description}
  \item [Damage:]  How much damage a bullet from this weapon will do to
other vehicles.  Note that the quality of their armor may reduce the damage 
they take from it.  See {\bf Armor}.
  \item [Range:]  How far the weapon can fire.
  \item [Ammo:]  How many bullets the weapon can hold.
  \item [Reload Time:]  After you have shot a weapon, how long it takes
to get the next shot off.
  \item [Ammo Speed:]  The speed of the weapon's bullets.
  \item [Ammo Cost:]  How much it costs to buy one bullet during a battle.
  \item [Heat:]  How much the heat of your vehicle will be increased every
time you fire this weapon.  See {\bf Heat Sinks}.
  \item [Weight:]  How much the weapon weighs.  Ammo has no weight.
  \item [Space:]  How much space the weapon takes up.
  \item [Cost:]  How much the weapon costs.  A full supply of ammo is
included free.
\end {description}

In addition to deciding what weapons to use, you must also determine where
to mount each weapon.  Mounting a weapon on a turret gives it the most
flexibility, but the turning rate of a turret is inversely proportional to
the weight of the stuff on it, so each additional weapon placed on a turret 
slows it down.

\begin{center}
{\footnotesize
\begin{tabular}{|c|r|r|r|r|r|r|r|r|r|r|}                            \hline
\multicolumn{11}{|c|}{\huge Weapons}                             \\ \hline
       &     &       &      & Reload & Ammo & Ammo & & & &      \\
Weapon & Dam & Range & Ammo & Time & Speed  & Cost & 
   Heat & Weight & Space & Cost                            \\ \hline \hline 

Light Machine Gun & 2 & 360 & 300 & 3 & 17 & 1 & 0 & 20 Kg & 200 L & 1000
\\ \hline
Machine Gun & 3 & 360 & 250 & 3 & 17 & 2 & 1 & 50 Kg & 225 L & 2200
\\ \hline
Heavy Machine Gun & 4 & 360 & 200 & 3 & 17 & 3 & 2 & 100 Kg & 250 L & 3000
\\ \hline
Light Pulse Rifle & 2 & 480 & 250 & 4 & 20 & 2 & 1 & 50 Kg & 200 L & 1500
\\ \hline
Pulse Rifle & 3 & 480 & 225 & 4 & 20 & 3 & 2 & 125 Kg & 225 L & 3300
\\ \hline
Heavy Pulse Rifle & 4 & 480 & 200 & 4 & 20 & 4 & 3 & 300 Kg & 250 L & 4500
\\ \hline
Light Autocannon & 3 & 400 & 250 & 3 & 22 & 4 & 2 & 200 Kg & 300 L & 3000
\\ \hline
Autocannon & 4 & 400 & 225 & 3 & 22 & 5 & 3 & 300 Kg & 350 L & 6000
\\ \hline
Heavy Autocannon & 5 & 400 & 200 & 3 & 22 & 6 & 4 & 500 Kg & 400 L & 10000
\\ \hline
Light Rkt Launcher & 4 & 600 & 150 & 7 & 25 & 8 & 3 & 600 Kg & 800 L & 7000
\\ \hline
Rkt Launcher & 6 & 600 & 125 & 7 & 25 & 9 & 5 & 900 Kg & 1200 L & 10000
\\ \hline
Heavy Rkt Launcher & 8 & 600 & 100 & 7 & 25 & 10 & 7 & 900 Kg & 1600 L & 15000
\\ \hline
Acid Sprayer & 6 & 160 & 100 & 3 & 10 & 10 & 0 & 300 Kg & 500 L & 3000
\\ \hline
Flame Thrower & 3 & 200 & 300 & 1 & 12 & 2 & 1 & 700 Kg & 500 L & 4000
\\ \hline
Mine Layer & 6 & 50 & 50 & 2 & 10 & 10 & 2 & 1000 Kg & 1000 L & 8000
\\ \hline
Heat Seeker & 8 & 1250 & 15 & 15 & 25 & 50 & 9 & 1000 Kg & 1800 L & 20000
\\ \hline
Oil Slick & 0 & 50 & 50 & 5 & 10 & 10 & 0 & 300 Kg & 500 L & 2000
\\ \hline
\end{tabular}
}
\end{center}

\newpage

\section{Armor}

\subsection*{How Armor Works}

Although there are many types of armor, you can have only one type
of armor on your vehicle.  Armor cost, weight, and space are proportional to
size of your vehicle's body, since a larger vehicle
has more area to cover with armor.
Each unit of armor provides one point of protection in one area of your 
vehicle.
The `hits off' of the armor determines how resistant the armor is to damage.
Vehicles armored with $x$ `hits off' take no damage from
weapons that do only $x$ damage, and sustain $x$ less damage from
better weapons.  When your armor on any side goes below zero, your vehicle
will explode.

\begin{description}
  \item [Example 1:]  Steel armor has 0 hits off, so if a bullet that does 5
points of damage hit a vehicle with steel armor, 5 points of steel armor
would be destroyed.

  \item [Example 2:]  Titanium armor has 2 hits off, so a bullet that does 5 
points of damage would only destroy 3 points of titanium armor.  The 2 
`hits off' negate 2 of the 5 damage points.

  \item [Example 3:]  Tungsten armor has 3 hits off, so a bullet that does 5 
points of damage would only destroy only 2 points of tungsten armor.  A 
bullet that does 3 points of damage would not affect tungsten armor at all.

\end{description}

\subsection*{What the Numbers Mean}

\begin{description}
  \item [Hits Off:]  Explained above.  The protection factor of the armor.
  \item [Weight Ratio:]  This multiplied by the size of your vehicle (see
{\bf Bodies}) will tell you how much each unit of armor will weigh.
  \item [Cost Ratio:]  This multiplied by the size of your vehicle (see
{\bf Bodies}) will tell you how much each unit of armor will cost.
\end{description}

\begin{center}
{\large
\begin{tabular}{|l|r|r|r|}                              \hline
\multicolumn{4}{|c|}{\huge Armor}                   \\ \hline
      & Hits & Weight & Cost  \\
Armor & Off  & Ratio  & Ratio                 \\ \hline \hline
Steel & 0 & 8 & 10                                   \\ \hline
Kevlar & 0 & 3 & 20                                  \\ \hline
Hardened Steel & 1 & 8 & 20                          \\ \hline
Composite & 1 & 4 & 30                               \\ \hline
Compound Steel & 2 & 8 & 40                          \\ \hline
Titanium & 2 & 5 & 70                                \\ \hline
Tungsten & 3 & 20 & 100                              \\ \hline
\end{tabular}
}
\end{center}

\newpage

\section{Specials}

Special abilities make your vehicle easier to use, and gives it a
wider variety of functions.  Each special is a separate
self-contained module capable of performing some useful function.

\begin{description}
  \item [Console:]  This module displays vital information about your vehicle
and its contents.  Don't leave home without it.

  \item [Mapper:]  This module remembers the locations
of all stationary objects in the maze that you have seen.
This allows you to plan routes through the maze efficiently, and also
helps you avoid getting cornered in dead ends.

  \item [Radar:]  This module detects all vehicles within 300 meters
and displays their approximate locations.  This not only helps to prevent
ambush, but allows you to track and chase down enemy vehicles.
Unfortunately, the radar is not foolproof, so occasionally vehicles will
slip by radar.  Also ghosting may occur (two blips for one vehicle).
\end{description}

\begin{center}
{\large
\begin{tabular}{|l|r|}    \hline
\multicolumn{2}{|c|}{\huge Specials}  \\ \hline
Special & Cost  \\ \hline \hline
Console & 250  \\ \hline
Mapper & 500   \\ \hline
Radar & 1000  \\ \hline
\end{tabular}
}
\end{center}

\section{Heat Sinks}

Many weapons have the annoying habit of creating heat in your
vehicle.  If this heat is not dissipated, you will not be able to fire
(when your heat reaches 100).

Each heat sink in your vehicle will reduce your vehicle's heat by one
every 5 frames (about 1/2 second during normal play).  Most vehicles have
at least one heat sink, usually 3 to 6.  If you plan on shooting many weapons
that generate lots of heat, consider buying 7 or more heat sinks.

Each heat sink weighs 500 Kg, takes up 1000 L of space, and costs 500.

\newpage

\section{Suspension}

\subsection*{How Suspension Works}

The cost of a standard suspension is included in the cost of the body.
If you want a better turning rate, you can use a heavy or active suspension
to improve your handling by 1 or 2 points.  This has the unfortunate
side effect of increasing the cost of the body.
On the other hand, if you don't need as much agility as the body has,
you can opt for a light suspension, saving on the cost of the body.

\subsection*{What the Numbers Mean}

\begin{description}
  \item [Handling Bonus:]  Add this to your vehicle's basic handling (see
{\bf Bodies}) to get the handling of your vehicle.
  \item [Cost Ratio:]  This multiplied by the size of your vehicle (see
{\bf Bodies}) will tell you how much it costs to use this kind of suspension.
\end{description}

\begin{center}
{\large
\begin{tabular}{|l|r|r|}            \hline
\multicolumn{3}{|c|}{\huge Suspension}   \\ \hline
           & Handling & Cost \\
Suspension & Bonus    & Ratio \\ \hline \hline
Light & -1 & 100    \\ \hline
Standard & 0 & 200  \\ \hline
Heavy & +1 & 400    \\ \hline
Active & +2 & 1000  \\ \hline
\end{tabular}
}
\end{center}

\section{Treads}

Wearing treads with better traction means that you will skid and slip less,
making it easier to make those tight turns.  However, because of the added
friction, your maximum speed will decrease.

\begin{description}
  \item [Friction:]  The percent of the maximum friction possible that
your vehicle has.
  \item [Cost Ratio:]  This multiplied by the size of your vehicle (see 
{\bf Bodies}) will tell you how much it costs to use these kind of treads.
\end{description}

\begin{center}
{\large
\begin{tabular}{|l|r|r|}    \hline
\multicolumn{3}{|c|}{\huge Treads}  \\ \hline
Treads & Friction & Cost Ratio    \\ \hline \hline
Smooth & .70 & 100  \\ \hline
Normal & .80 & 200  \\ \hline
Chained & .90 & 400 \\ \hline
Spiked & 1.00 & 1000 \\ \hline
\end{tabular}
}
\end{center}

\newpage

\section{Bumpers}

If you wear better bumpers, you will rebound from obstacles (such as walls
and other vehicles) faster, and also take less damage than usual. 

\begin{description}
  \item [Elasticity:]  A general measure of the elasticity of your bumpers.
  \item [Cost Ratio:]  This multiplied by the size of your vehicle (see 
{\bf Bodies}) will tell you how much it costs to use these kind of bumpers.
\end{description}

\begin{center}
{\large
\begin{tabular}{|l|r|r|}  \hline
\multicolumn{3} {|c|}{\huge Bumpers}  \\ \hline
Bumpers & Elasticity & Cost Ratio    \\ \hline \hline
None & 0 & 0       \\ \hline
Normal & .05 & 200 \\ \hline
Rubber & .15 & 400 \\ \hline
Retro & .25 & 1000  \\ \hline
\end{tabular}
}
\end{center}

\section{Formulas}

Herein lie all the formulas for calculating all aspects of your vehicle.
In general, you don't need to worry about them; Xtank will do all of the
work for you.  They are included for the curious.

\begin{center}
\(
\mbox{Total weight} = \mbox{body weight} + \mbox{engine weight} 
  + \mbox{total armor} \times \mbox{armor cost ratio} \times
    \mbox{body size} \) \\ \(
  + \mbox{heat sinks} \times \mbox{heat sink weight}
  + \mbox{total weapon weight}
\) \\ 
\end{center}

\begin{center}
\(
\mbox{Total space} = \mbox{engine space} + 
  + \mbox{total armor} \times 3 \times
    \mbox{body size} \) \\ \(
  + \mbox{heat sinks} \times \mbox{heat sink space} 
  + \mbox{total weapon space}
\) \\ 
\end{center}

\begin{center}
\(
\mbox{Total cost} = \mbox{body cost} + \mbox{engine cost} 
  + \mbox{heat sinks} \times \mbox{heat sink cost} \) \\ \(
  + \mbox{total weapon cost} + \mbox{total special cost} 
  + (\mbox{total armor} \times \mbox{armor cost ratio} \) \\ \(
  + \mbox{suspension cost ratio} + \mbox{tread cost ratio} 
  + \mbox{bumper cost ratio}) \times \mbox{body size}
\) \\ 
\end{center}

\begin{center}
\(
\mbox{Handling} = \mbox{body handling base} 
                    + \mbox{suspension handling adj}
\) \\ 
\end{center}

\begin{center}
\(
\mbox{Max. speed} = {\frac{\sqrt[3]{\frac{\mbox{engine power}}
                       {\mbox{body drag}}}}{\mbox{tread friction}}}
\) \\ 
\end{center}

\begin{center}
\(
\mbox{Acceleration} = 16 \times {\frac{\mbox{engine power}}
                         {\mbox{total weight}}}
\) \\
\end{center}

\end{document}




